---
title: "p8105_hw2_sz2800"
author: "Stephanie Zhen"
date: "10/04/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(tidyverse)
```

### Question 1: Trash Wheel

Round the number of sports calls to the nearest integer and converts the result to an integer variable (using as.integer)
```{r}
dumpster_df = read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel", skip =1) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = as.integer(sports_balls)
  )
```

2017 Precipitation import + adding year
```{r}
trash17_df = read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet =  "2017 Precipitation", skip =1, n_max = 12) %>%
  janitor::clean_names() %>%
    mutate(
      year = 2017)
```

2018 Precipitation import + adding year
```{r}
trash18_df = read_excel("./Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet =  "2018 Precipitation", skip =1, n_max = 12) %>%
  janitor::clean_names() %>%
    mutate(
      year = 2018)
```

Joining 2017 and 2018 datasets together. Change month name by using month.name
```{r}
prec_join = 
  full_join(trash17_df, trash18_df) %>%
  mutate(
    month = month.name[month]
  )
```
The number of observations in the joined precipitation 2017 and 2018: `r nrow(prec_join)`

The key variables in prec_join and dumpster_df: `r colnames(prec_join)`

The total amount of precipitation in 2018:`r sum(pull(trash18_df, total))`

Median sports balls in 2017:`r filter(dumpster_df, year == "2017") %>% pull(sports_balls) %>% median()`


### Question 2: fivethirtyeight_datasets

pols_month
```{r}
pols_clean = read_csv("./fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day"), convert = TRUE)
pols_clean_tidy = 
  pols_clean %>%
  mutate(month = month.abb[month]) %>%
  mutate(president = ifelse(prez_gop > 0, "gop", "dem")) %>%
  select(-day, -prez_gop, -prez_dem)
```

snp: clean up
```{r}
snp_clean = read_csv("./fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("month", "day", "year"), convert = TRUE) %>%
  mutate(month = month.abb[month]) %>%
  select(year, month, close)
```
umemployment_clean wide to long
```{r}
unemp_clean = read_csv("./fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(
    cols = Jan:Dec,
    names_to = "month",
    values_to = "unemp_perc"
  ) %>%
  select(year = Year, month, unemp_perc) 
```
Joining/merging dataset: Left joined pols_clean_tidy, snp_clean then left join again with resulting dataset and unemp_clean
```{r}
pols_snp_join= 
  left_join(pols_clean_tidy, snp_clean, by = c("year", "month")) 

unemp_p_s_join =
   left_join(pols_snp_join, unemp_clean, by = c("year", "month")) 
```

### Popular Baby Names dataset
cleaning baby names dataset
```{r}
p_baby_clean = read_csv("./Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  rename("name" = "childs_first_name", "year" = "year_of_birth") %>%
  mutate(ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER", "BLACK NON HISP" = "BLACK NON HISPANIC", "WHITE NON HISP" = "WHITE NON HISPANIC")) %>%
  mutate(name = str_to_upper(name)) %>%
  distinct()
```

Ranks of the name: "OLIVIA" over time.
```{r}
olivia_df = p_baby_clean %>%
  filter(name == "OLIVIA", gender == "FEMALE") %>%
  select(name, year, ethnicity, rank) %>%
  pivot_wider(
    names_from = year,
    values_from = rank
  ) %>%
    knitr::kable()
```

Most popular, rank = 1, male birth name over time.
```{r}
pop_male_df = p_baby_clean %>%
  filter(rank == 1, gender == "MALE") %>%
  select(name, year, ethnicity) %>%
  pivot_wider(
    names_from = year,
    values_from = name
  ) %>%
    knitr::kable()
```

ggplot:
```{r}
male16_df = p_baby_clean %>%
  filter (gender == "MALE", year == 2016, ethnicity == "WHITE NON HISPANIC")

male16_plot = ggplot(male16_df, aes(x = rank, y = count)) +
  geom_point()
```






